---
title: "HW 2"
author: "Sam Loewen"
date: "9/27/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

## Problem 1

First I will read and clean the Mr. Trash Wheel sheet:

*   specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
*   use reasonable variable names
*   omit rows that do not include dumpster-specific data
*   round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

```{r}
trash_wheel = 
  read_excel("./data/HealthyHarborWaterWheel.xlsx", sheet=1)
  skimr::skim(trash_wheel)

#cleaning names
trash_wheel = 
  janitor::clean_names(trash_wheel)
  names(trash_wheel)
 
#removing 'total' rows, x15 column   
trash_wheel=
  tidyr::drop_na(trash_wheel, dumpster)

#removing x15 column, rounding sports balls
trash_wheel=
  select(trash_wheel,-x15)
  mutate(trash_wheel, sports_balls = as.integer(sports_balls))
```






