---
title: "HW 2"
author: "Sam Loewen"
date: "9/27/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(knitr)
```

## Problem 1

First I will read and clean the Mr. Trash Wheel sheet:

*   specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
*   use reasonable variable names
*   omit rows that do not include dumpster-specific data
*   round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

```{r}
trash_wheel = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  select(-x15, -x16, -x17) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

Next I will read and clean precipitation data for 2017 and 2018. For each omit rows without precipitation data and add a variable year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

```{r}
precip_17 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 6, skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = 2017) %>% 
  select(year, month, total)

precip_18 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = 2018) %>% 
  select(year, month, total)

precip_all = 
  full_join(precip_17, precip_18) %>% 
  mutate(month = month.name[month])
```

```{r include = FALSE}
tw_sb =
  trash_wheel %>%
  filter(year == 2017) %>%
  select(year, sports_balls)
```

My trash_wheel data set has `r nrow(trash_wheel)` observations, and my precip_all data set has `r nrow(precip_all)` observations. Trash_wheel includes variables to show the number of items removed from the harbor, such as `glass_bottles` and `grocery_bags`. We can also see that the median number of `sports_balls` in the dumpster in 2017 was `r median(pull(filter(trash_wheel, year==2017), sports_balls))`. The precip_all dataframe can show us that the total precipitation in 2018 was `r sum(pull(precip_18, total))` inches.

## Problem 2

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r}
pols =
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day")) %>%
  mutate(prez = ifelse(prez_gop == 1, "gop", "dem")) %>% 
  mutate(year = as.numeric(year), 
         month = replace(month, month == "01", "jan"),
         month = replace(month, month == "02", "feb"),
         month = replace(month, month == "03", "mar"),
         month = replace(month, month == "04", "apr"),
         month = replace(month, month == "05", "may"),
         month = replace(month, month == "06", "jun"),
         month = replace(month, month == "07", "july"),
         month = replace(month, month == "08", "aug"),
         month = replace(month, month == "09", "sep"),
         month = replace(month, month == "10", "oct"),
         month = replace(month, month == "11", "nov"),
         month = replace(month, month == "12", "dec"))%>% 
  select (-prez_gop, -prez_dem, -day)
```

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
snp =
  read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year")) %>% 
  mutate(year = as.numeric(year)) %>% 
  select (year, month, -day, close) %>% 
  arrange (year, month)
```

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r}
unemp=
  read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec, 
    names_to = "month", 
    values_to = "snp")
```

Now joining the datasets.

```{r}
q2_merge =
  left_join(pols, snp, by = c("year","month"))

final_db = 
  left_join(q2_merge,unemp, by =c("year","month"))
```

These datasets provide information on who was in power by month starting in the late 1940s and early 50s including `snp`, which provides data on the S&P 500, and unemployment data. The resulting database `final_db` combines all these datasets to show variables such as number of senators per party (`sen_dem` and `sen_gop`), the political party of the president (`prez`) and `snp`, and unemployment data. The databases has `r nrow(final_db)` observations and covers January 1947 to June 2015.

## Problem 3

Load and tidy the data. Note that, although these data may seem fairly well formatted initially, the names of a categorical predictor and the case structure of string variables changed over time; you’ll need to address this in your data cleaning. Also, some rows seem duplicated, and these will need to be removed (hint: google something like “dplyr remove duplicate rows” to get started).

```{r}
baby_names = 
  read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"),
    ethnicity = replace(ethnicity, ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC"),
    ethnicity = replace(ethnicity, ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC"), 
    childs_first_name = toupper(childs_first_name)) %>% 
  distinct()
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. Produce a similar table showing the most popular name among male children over time.

```{r}
#olivia
baby_names_o =
  baby_names %>%  
  filter(childs_first_name == "OLIVIA") %>% 
  select (-gender, -childs_first_name, -count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") %>% 
  knitr::kable()

baby_names_o
```

```{r}
#boy names
baby_names_boy = 
  baby_names %>%
  filter(gender == "MALE", rank == 1) %>% 
  select (-gender, -count, -rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name") %>% 
  knitr::kable()

baby_names_boy
```

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
scatter_names =
  baby_names %>% 
  filter (gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016)

scatter_names %>% 
  ggplot(aes(x = rank, y = count)) + 
  geom_point() +
    labs(title = "Rank v count of names, among non-hispanic white boys born in 2016", 
    x = "Rank", 
    y = "Count")
```
